---
title: Information for Outgoing Students
---

::: callout-note
## Information on available places and partner universities of the FAU

The interactive map and table below only show the **partner universities of the Chair for Communication Science**. You can find an overview of all the places abroad offered by FAU at:

🔗 [Partner universities for ERASMUS+](https://www.fau.de/education/international/wege-ins-ausland/studieren-im-ausland/erasmus-studium/erasmus-partnerhochschulen/) 

🌍 [Interactive world map](https://www.fau.eu/education/international/going-abroad/studying-abroad/)
:::

## Partner universities

```{r session-setup}
#| echo: false
#| message: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
    readr, qs,
    here, tidyverse, 
    leaflet, raster, 
    osmdata, osmextract,
    glue,
    gt
)
```

```{r optional-create-new-qs-data}
#| echo: false
#| eval: false
source(here("R/data-creation.r"))
```

```{r import-db}
#| echo: false
#| message: false
df_en <- qread(here("data/universities-geo_locations_en.qs"))
```

```{r create-icons}
#| echo: false
#| column: page
#| fig-align: center
icons <- awesomeIcons(
    markerColor = "darkblue"
)

df_en <- df_en %>%
  separate(lat_long, into = c("lat", "lng"), sep = ",", convert = TRUE)

leaflet(df_en) %>%
  addTiles() %>%
  addAwesomeMarkers(lng = ~lng, lat = ~lat, popup = ~info, icon = icons)
```

```{r create-icons-troubleshooting}
#| echo: false
#| eval: false
invalid_rows <- which(!validUTF8(df_en$info))

# Zeige problematische Zeilen an
df_en[invalid_rows, ]
df_en$info <- iconv(df_en$info, from = "unknown", to = "UTF-8", sub = "byte")
unique(Encoding(df_en$info))
# Finde Zeilen mit "unknown" Kodierung
unknown_encoding_rows <- which(Encoding(df_en$info) == "unknown")

# Zeige die betroffenen Zeilen an
df_en[unknown_encoding_rows, ]
df_en$info <- iconv(df_en$info, from = "windows-1252", to = "UTF-8", sub = "byte")
```

```{r create-table}
#| echo: false
#| column: page

library(gt)

df_en %>%
    mutate(
        duration = paste(duration, "months", sep = " "),
        uni_link = glue("[{uni}]({link})"),
        uni_link = map(uni_link, gt::md)
    ) %>% 
    dplyr::select(country, uni_link, places:languages, wise:sose) %>% 
    gt::gt(id = "table_2") %>% 
    cols_label(
        country = "country",
        uni_link = "University", 
        places = "places",
        duration = "duration of exchange",
        languages = "Courses available in",
        wise = "winter term",
        sose = "summer term"
    ) %>% 
    gtExtras::gt_highlight_rows(
        rows = c(7, 10, 12),
        fill = "#18B4F1", 
        alpha = 0.075,
        font_weight = "normal",
        bold_target_only = FALSE
    ) %>%
    tab_footnote(
    footnote = "only BA students",
    placement = "right", 
    locations = cells_body(
      columns = 3, row = c(7, 10, 12))
    ) %>% 
    gtExtras::gt_theme_538()

```

## Requirements for participation

::: callout-important
## Important notes

-   Please note the general [eligibility requirements](https://ib.wiso.fau.de/en/outgo/application/) for the exchange program on the website of the [International Relations Office](https://ib.wiso.fau.de/en/).
-   The programs of the Chair of Communication Science are **exclusively social science exchange programs!**\
:::

### Before the stay:

::: callout-warning
#### Important! Please note the following:

-   Successful completion of **at least one module in Communication Science** ([Chair](https://www.commsci.rw.fau.eu/teaching/)). Applies **to both Bachelor's and Master's students**. **For *Bachelor* students, we particularly recommend attending one of our lectures**:
    -   [📖 Medien und Kommunikation](https://www.kowi.rw.fau.de/lehre/bachelor/medien-und-kommunikation/)
    -   [📖 Digital Technologies & Society](https://www.kowi.rw.fau.de/lehre/bachelor/digital-technologies-and-society/)
-   Applicants should also have **sufficient knowledge of the respective national language or a good command of English**.
:::

### During the stay:

-   Successful completion of at least one communication science course at the specific institution (faculty, department, etc.) of the respective partner university with which the cooperation agreement exists (see list of cooperating universities).
-   In addition, the overall focus of the courses taken should be on communication science or social science courses from the program of the specific partner institution. Taking courses at other faculties / institutions / departments of the partner university is only possible in occasional cases or in some cases not at all!

### After the stay:

-   After your stay abroad, there are still a few steps to take in order to have your credits entered and converted. You can find helpful information on this in the section [Accreditation of academic achievements].

## Application process

The application for a place at one of our partner universities is coordinated by the [International Relations Office](https://ib.wiso.fau.de/en/). There you will also find detailed [information on the application process](https://ib.wiso.fau.de/en/outgo/application/).

::: callout-note
## Required application documents

<!-- FIXED Auflistung Bewerbungsunterlagen (z.B. Notenauszug, CV, Motivationsschreiben) und Bewerbungsfrist für „Erste-Runde“ Bewerbungen für Aufenthalte im Wintersemester und Sommersemester sowie für eventuelle Restplätze (dann vorwiegend für Sommersemester). Empfehlung für die fachinterne Bewerbungsfrist ist Mitte Dezember. Kommentar zum Hintergrund für Sie (nicht für Studierende): die Frist für Meldung (sog. „Nominierung“) an S-International ist aktuell Mitte Februar und wird möglicherweise in den kommenden Jahren auf Mitte Januar vorverlegt aufgrund der frühen Mittelanfragefrist bei der Nationalen Erasmus Agentur. -->

The following documents must be uploaded via the [online portal](https://ib.wiso.fau.de/outgo/bewerbung/bewerbungsformular-outgoing/) of the [International Relations Office](https://ib.wiso.fau.de/en/) in PDF format (max. 1MB per document):

-   Current certificate of enrollment
-   Letter of motivation (in German or English, maximum 1 A4 page)
-   Curriculum vitae in tabular form (in German or English, maximum 2 A4 pages)
-   Grades documents:
    -   for BA applicants, grade transcript from [MeinCampus](https://www.campus.uni-erlangen.de/qisserver/pages/cs/sys/portal/hisinoneStartPage.faces),
    -   for MA applicants, a German BA transcript or copy of the international BA transcript (official conversion is transferred internally according to the so-called Eignungsfeststellungsnote of the FAU admission office).
-   Language certificates (see \[[Language Requirements]{.underline}\])(https://ib.wiso.fau.de/en/outgo/preparations/language-skills-tests/)
-   Proof of internships, social or volunteer work (maximum 2 proofs)
:::

## Accreditation of academic achievements

For accreditation the academic achievements must meet the requirements of the chair in terms of content and workload. As a guideline, the general information about the contents of the communication science courses can be found in the [module handbooks](https://www.wiso.rw.fau.eu/study/current-students/module-handbooks/).

Further helpful information (e.g., regarding grade transfer or the respective contact persons of the individual departments) can be found at:

🔗 [Information on Accreditation and grade transfer](https://ib.wiso.fau.de/en/outgo/nach-dem-aufenthalt/accreditation-grade-transfer/)

🔗 [Contacts of the individual departments at WiSo](https://ib.wiso.fau.de/en/outgo/nach-dem-aufenthalt/contacts-accreditation-wiso/)

### Learning Agreements

<!-- FIXME Angabe zuständige/r Erasmus Departmentkoordinator/in -->

<!-- FIXME Angabe Unterschriftsbefugte/r für Learning Agreement for Studies (DLA) -->

<!-- FIXME Beschreibung des Prozesses der Anerkennung von Leistungen aus dem Ausland (oder Verlinkung auf dementsprechende Webseite der Fakultät) und Benennung Anerkennungsbeauftragte/r -->

<!-- FIXME Benennung von Beschwerdestelle (ggf .Departmentleitung) bei Kritik an Auswahl oder Anerkennungsprozess, ergänzende Verlinkung hierher: https://www.fau.de/fau/willkommen-an-der-fau/internationalisierung/erasmus/#collapse_35  -->

As a general rule, the student has to agree with the departmental supervising chair on the lecture programme necessary for recognition, if possible before the start or at the latest at the beginning of the stay abroad, as far as this is possible on the basis of the information available at the host university.

It is therefore advisable to think about what is to be studied at the partner university already in the application phase. To secure this agreement, a [Learning Agreement](https://ib.wiso.fau.de/en/outgo/preparations/learning-agreements/) is obligatory.

Please submit requests for Learning Agreements via [E-Mail](mailto:wiso-erasmus-kowi@fau.de), providing the following information in full:

-   Completed Learning Agreement (Please use the [current form](https://ib.wiso.fau.de/en/outgo/preparations/learning-agreements/) of the International Relations Office)

-   The communication science courses you have already attended

-   Web link to the course description of the course proposed for recognition at the partner university (this should include course title, content and scope, assessment, number of ECTS).

### Notes for Bachelor's students

A study abroad is particularly recommended for students who choose Communication Studies as their specialisation. Although it is generally possible to take the basic courses abroad, they must correspond to the courses offered by the department and have an introductory character. Since the course offerings abroad are often limited, credit transfer for basic courses cannot be guaranteed.

```{=html}
<!-- TODO Check infos for: Praxisaufenthalte Studierende Erasmus (Optional) --\> 

<!-- TODO Specifically for: Outgoing (kein Muss, aber herzliche Bitte, dies bekannt zu machen) -->
```

<!-- FIXME Hinweis, dass Praxisaufenthalte eine Alternative sein können zu einem Semesteraufenthalt zu Studienzwecken, da sie auch kürzer möglich sind und die sog. „Employability“ steigern; und Verlinkung: https://www.fau.de/education/international/wege-ins-ausland/praxisaufenthalt-im-ausland/  -->

<!-- FIXME Hinweis: unter bestimmten Voraussetzungen kann man Erasmus Praktikums-Stipendien für Praktika und praktische Teile von Abschlussarbeiten beantragen, Verlinkung: https://www.fau.de/education/international/wege-ins-ausland/praxisaufenthalt-im-ausland/erasmus-praktika/  -->

<!-- FIXME Nennung Unterschriftbefugte/r Learning agreement for Traineeship, und auch ggf., wenn vorhanden, Praktikumsamt oder Berater/in -->

<!-- FIXME Infos über potentielle Anerkennung als Praxis-Modul, Abschlussarbeit oder als Soft skills, falls zutreffend bzw. in Prüfungsordnung vorgesehen -->

## Linguistic and intercultural preparation

<!-- TODO Specifically for: Sprachliche und interkulturelle Vorbereitung (Optional aber sehr wünschenswert) -->

<!-- FIXED Hinweis auf Wichtigkeit und Nützlichkeit der rechtzeitigen Vorbereitung auf den Auslandsaufenthalt und die Optionen, die die FAU dafür beitet, z.B. Hinweis auf Kurse Schlüsselqualifikationen (Interkulturelle Trainings z.B. des Ziwis oder auch im Sprachenzentrum (über campo auffindbar) https://www.ziwis.fau.de/lehre/schluesselqualifikationen/#collapse_1  -->

<!-- FIXED Hinweis auf Sprachkurse Sprachenzentrum: https://sz.fau.de/sprachkurse/  -->

<!-- FIXED Hinweis auf Erwerb Zertifikat Interkulturelle Kompetenz für Studierende: https://www.ziwis.fau.de/zertifikat-interkulturelle-kompetenz-fuer-studierende-zertifikationsprogamm/  -->

<!-- FIXED Verlinkung zu Buddy Programm: https://www.fau.eu/education/international/from-abroad/welcome-programme-for-international-students/buddy-programme/  -->

In time preparations for studying abroad are a part of the exchange process that should not be underestimated. FAU offers many useful options, e.g.,

-   Courses on specific topics (e.g., intercultural training at [ZIWIS]((https://www.ziwis.fau.eu/home/courses/key-qualifications/)) or at the Language Centre (can be found via campo))
-   [Language Courses](https://sz.fau.eu/courses/) offered by the Language Centre
-   Acquisition of the certificate [*Intercultural Competence for Students*](https://www.ziwis.fau.de/zertifikat-interkulturelle-kompetenz-fuer-studierende-zertifikationsprogamm/)
-   [Buddy programme](https://www.fau.eu/education/international/from-abroad/welcome-programme-for-international-students/buddy-programme/)